---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
header-includes:
  -  \usepackage{hyperref}
biblio-style: apsr
title: "Working from home: estimations without surveys"
thanks: "TBD"
author:
- name: Henning Holgersen
  affiliation: Statistics Norway
abstract: "After the Covid-19 pandemic outbreak, working from home has become the norm for many. Surveys have tried to capture the extent to which people actually work from home, which is a type of problem surveys are good at. We have attempted to use alternative approaches and data sources to say something about who might be working from home. Alternative data sources are rarely a drop-in replacement for surveys. Instead, they answer different but related questions."
keywords: "big data, surveys, work, text"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,
                      message=FALSE, warning=FALSE,
                      fig.path='figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })

library(tidyverse)
```


# Introduction

The Covid-19 pandemic has led practically anyone who can to work from home. A priori, we didn't have a lot of information about what percentage of the workforce actually had this opportunity, and we knew less about what characterizes these people. It is easy to guess that office jobs generally can be done from home, but an office job is not clearly defined in official statistics.

Surveys after the pandemic took effect have shown that more than 40 percent of the workforce now works from home. In contrast, some surveys prior to the pandemic found that ~30 percent of the workforce were permittet to work from home at least on occasions. Naturally, the two questions are different. During the pandemic, practically all employers will let employees work from home if possible. Before, working from home was somewhat of a perk for the employer to grant if they had the opportunity and were so inclined.

In these situations, there are no obvious alternatives to surveys. Two possibilities are evaluated here: Full text job ads, and annotating the ISCO taxonomy by interpreting the documentation.

Job advertisements may mention in their ad if they have good options for working from home. Working from home at times is a valuable perk for many, and employers therefore may have an incentive to mention it. This approach is contingent of job advertisements actually being available to analyze, and it is very difficult to say anything about the total number of remote-enabled jobs out there. Even saying something about the relative frequency of remote jobs between industries, occupations etc. requires calibration. But job advertisements are a rich source of data, and if they are readily available, it would be extravagant to dismiss them.

The ISCO taxonomy does not involve itself with wether or not the job can be done remotely. The documentation, however, includes short but informative descriptions of the job. Humans reading these descriptions will often have an immediate understanding of the possibilities for doing this job remotely. In order to classify each occupation as remote-friendly or not. This is a fairly simple job to do, and the ISCO documentation is available for anyone online. For anyone who already has register or survey data which includes ISCO codes, this is a simple approach to augment the data they already have.

The four data sources above (two surveys and to alternative sources), answer four different but related questions:

- The surveys before covid asked employees if they had the *opportunity* to work from home, and if so how often.
- The surveys after covid primarily ask employees *if* they worke from home (although some nuances exist)
- The job ads can tell us about job openings the employer advertises as remote friendly
- The ISCO annotations can tell if there is anything about the job itself that prevents it from being performed from home.

These differences can either be vital and not to be confused, relatively minor nuances, or informative in themselves.



# What the surveys say




# Annotating the ISCO standard

The ISCO standard organizes jobs into a set of groups according to the tasks and duties undertaken in the job. Using the detailed task descriptions listed in the ISCO-08 documentation, we try to provide a assertion of whether an occupation is likely able to be performed from home. To do this, we created a public labeling job through Amazon Mechanical Turk [@Turk2020]. Each occupation was presented together with a brief description. The exact question formulation was “Can this type of job likely be performed from a home office?”, and an example of a job description could be:

>Social work and counselling professionals provide advice and guidance to individuals, families, groups, communities and organizations in response to social and personal difficulties. They assist clients to develop skills and access resources and support services needed to respond to issues arising from unemployment, poverty, disability, addiction, criminal and delinquent behaviour, marital and other problems.

The respondent was asked to evaluate wether it was likely that the job could be performed primarily from a private home. The alternatives were “Yes”, “No” and “Unknown”, which were provided with the following description:

1. *Yes: This job can be performed primarily from an office in a private home*

2. *No: Substatantial parts of this job must be performed outside the employees home*

3. *Unknown: There is not enough information to decide*

In order to reduce the serendipity in the labels, we acquired five labels
from different respondents for each occupation, and we provided an
`uncertain` option in addition to the `yes/no` options in order to
reduce arbitrary responses to uninformative occupation descriptions. The final labels include an uncertainty measure which shows that some of the occupations were evaluated differently by different annotators, but no occupation was given a final label of "Unknown" which means we can treat the remote-friendly annotation as a binary variable.

Since the job was on Mechanical Turk, there respondents were not subject matter experts, and likely reside in different countries. This adds to the importance of obtaining more than one label per occupation, but the number of labels does not correct for possible cultural differences - it is possible that some jobs that cannot be performed remotely in other countries can be performed remotely in Norway. We should consider the annotations as `international`, which is also true for the ISCO-08 standard itself.

There are two additional approaches to the annotation approach: The ISCO documentation includes examples of typical tasks, which can be categorized on their own, and an aggregate measure of "what fraction of this job can be done at home" can be calculated. Of course, this measure will not be able to take into account the amount of time spent at each task - there may be no substitute for surveys here.

The other approach, very similar, is to use ontologies like ESCO, which includes tasks. This may open the possibility of calculating "nearest remote neighbor", building on similarities between 

# Utilizing job ads

The market for job ads is relatively concentrated in Norway. For over 10 years, there has been two major actors in this market: the commercial website finn.no, and the norwegian welfare administration, nav.no. Many ads have been posted on both sites, but starting in 2018 the nav.no website republishes ads from finn.no. In the summer of 2019, NAV published an archive of job ads going back to 2002. This data set (one giant csv file) contains the title and full text of the ad, information about the employer including organizational number (ID from the central business register), and an ISCO classification made by employees at NAV.

The job ads can tell us if the employer explicitly advertise rempte possibilities, but we are faced with a question: How do we know? There are basically three approaches.

Searching for specific words or phrases that are highly suggestive of remote possibilities. This approach works if the language, like norwegian, has a small number of words that indicate this. The solution is low-tech, but may be surprisingly effective for certain languages.

The search can be amended by using word embeddings to find similar words. Simple word embeddings, however, will not allow us to find phrases that indicate remote possibilities. In order to use that approach, advanced NLP techniques such as Named Entity Recognition must be considered.

Since results are highly language dependent, we will limit our foray into NLP.

Using a simple search, we find 3500 mentions of *hjemmekontor* or *heimekontor*. This is very low considering there are 2.6 million ads, but still enough to make some inferences.

```{r}
#load('wfh.RData')
loadd(wfh_plot_data)

wfh_plot_data %>% 
  ggplot(aes(aar, andel_stillinger)) +
  geom_smooth()
```

That's all folks!

```{r}
head(iris)
```


